Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-03-05 doc/paper.Rmd
changed some of the word formations

Markings made by Screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length. The Ruger pistol barrels used in the  Hamby study have a 9mm caliber. This means that bullets used in these pistols have land markings of almost $(1/4)^{th}$ the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising a problem of distinctiveness. One way to investigate this is to evaluate bullet striations at different resolutions. This gives us an incentive to use both the NIST and Iowa State scans and investigate the effect of length of the digitized mark on the working of the algorithm. Apart from this, the algorithm developed by \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature, whereas striations on bullets are made on their curved surfaces. Therefore, using methods proposed for toolmarks need adaptation in order to give tangible results for bullets. Another aspect of investigation is the performance of the algorithm for different levels of smoothing when signatures are being compared as opposed to when raw markings or profiles are being compared.

The Chumbley method and algorithm works on the premise of comparison windows which are supposed to represent small segments of pixels of equal sizes in the two markings that are to be compared in a predetermined fashion. They term them as windows of optimization and validation. The optimization window is selected in the first step called the optimization step where areas/windows of best agreement between the the two markings are found.

The window of optimization for bullet striations are shorter as bullet signatures in terms of absolute length in mm, are smaller compared to toolmarks. Although in terms of digitzed pixels there is no straightforward way of establishing what window size is to be chosen. One important way to address this is to keep the number of windows of optimization sufficiently large or not too small. This means, if the number of windows are too small we inadvertently use bigger window sizes as we want to compare all segements of one mark with the segments of the other mark.This leads to not giving enough emphasis to smaller segements of the two marking that might be in high agreement. In terms of absolute length we end up comparing smaller segments between the two markings while dealing with bullet lands, than those being compared for toolmarks.

>>>>

Markings made by Screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length. The Ruger pistol barrels used in the  Hamby study have a 9mm caliber. This means that bullets used in these pistols have land markings of almost $(1/4)^{th}$ the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising a problem of distinctiveness. One way to investigate this is to evaluate bullet striations at different resolutions. This gives us an incentive to use both the NIST and Iowa State scans and investigate the effect of length of the digitized mark on the working of the algorithm. Apart from this, the algorithm developed by \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature, whereas striations on bullets are made on their curved surfaces. Therefore, using methods proposed for toolmarks need adaptation in order to give tangible results for bullets. Another challenge that arises from these topographical differences, is the understanding of that effect of different levels of smoothing. Especially investigating its effect on the algorithm performance when signatures are being compared and when raw markings or profiles are being compared.

Some other challenges associated with the adaptation process call for a brief understanding of the methodology involved. The Chumbley method and algorithm works on the premise of comparison windows that are supposed to represent small segments of pixels of equal sizes in the two markings.  These windows are to be compared in a predetermined fashion and are termed as windows of optimization and validation. The optimization window is selected in the first step called the optimization step where areas/windows of best agreement between the the two markings are found.

The window of optimization for bullet striations are shorter as bullet signatures in terms of absolute length (in mm), are smaller compared to toolmarks. Even in terms of digitized pixels, there is no straightforward way of establishing what window size is to be chosen. One important way to address this is to keep the number of windows of optimization sufficiently large or not too small. This means, if the number of windows are too small we inadvertently use bigger window sizes as we want to compare all segements of one mark with the segments of the other mark. This leads to not giving enough emphasis to smaller segements of the two marking that might be in high agreement. In terms of absolute length we end up comparing smaller segments between the two markings while dealing with bullet lands, than those being compared for toolmarks.
